Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-04-14 paper.tex
revisions

% Created 2017-04-13 Thu 18:06
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org0a430e5}

@@

\section{Background}
\label{sec:orgc5b76aa}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgce77dd1}

@@

\subsection{Passthought authentication}
\label{sec:orgddf4083}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org7c65df9}

@@

\section{User attitudes and perceptions}
\label{sec:org1e82386}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org6b82b47}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orgb912760}

@@

\label{fig:rank}
\begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{./figures/rankings.png}
\caption{``Please rank the following sensors in how likely you believe they are to reveal what a person is thinking and feeling.'' Mean Likert responses (Not at all\ldots{}Very informative). Higher bars indicate higher rank, or more likely to reveal what one is thinking or feeling.''}
\end{figure*}

\section{Diversity and security of passthoughts}
\label{sec:orga4f01a7}

@@

\section{Two studies on passthoughts}
\label{sec:orgda992dc}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org647a2bd}

@@

\subsubsection{Study protocol}
\label{sec:org5f58e5b}

@@

\subsubsection{The effect of feedback}
\label{sec:org68a3874}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org236c4c5}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgdb1427a}

@@

\subsubsection{Study protocol}
\label{sec:org0ed08a7}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:org5eb9cdc}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org40f3dce}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orgd44d815}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org7e5111c}

@@

\subsection{Privacy}
\label{sec:org8a9010b}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org22dc2d3}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org9627bbb}

@@

\section{Further Future Directions}
\label{sec:orgd918e94}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org80cf40b}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org7d58379}

@@

\subsection{Organic passwords}
\label{sec:orgc942193}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Neuroscience of authentication}
\label{sec:orgda2b5f8}

Where authenticity is nominally concerned with proving that you are who you say you are,
a less-frequently-asked question in the authentication literature is,
``are you really yourself?''
We all sometimes do or say regrettable things when we are feeling ``not quite ourselves,'' sometimes using devices or services with which we have authenticated ourself.
Can authentication ever verify not only your posession of your body, but of your ``right mind''?

A question raised earlier surrounds wehre passthoughts could still work if a person is drunk, having a migraine, or in distress (Section 3). 
Even if passthoughts fails when a user is in such an ``off-baseline'' state, 
passthoughts still may have utility (perhaps even \emph{added} utility) in certain authentication contexts.
For example, one may wish to allow themselves access to certain resources (e.g. bank accounts) when one's resting EEG state is not too much different from a pre-recorded baseline.

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as I am aware, novel, and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this still-novel method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org495f530}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org70dd836}

@@

\section{Conclusion}
\label{sec:orga01fe32}

>>>>

% Created 2017-04-14 Fri 09:02
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org2a2b726}

@@

\section{Background}
\label{sec:org246c216}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgda4a571}

@@

\subsection{Passthought authentication}
\label{sec:org1eb6c99}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org7a20391}

@@

\section{User attitudes and perceptions}
\label{sec:org6284e0b}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org605033b}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org1129bb0}

@@

\label{fig:rank}
\begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{./figures/rankings.png}
\caption{``Please rank the following sensors in how likely you believe they are to reveal what a person is thinking and feeling.'' Higher bars indicate higher rank, or more likely to reveal what one is thinking or feeling.''}
\end{figure*}

\section{Diversity and security of passthoughts}
\label{sec:orgdd2553c}

@@

\section{Two studies on passthoughts}
\label{sec:org78fbfbd}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org08c501f}

@@

\subsubsection{Study protocol}
\label{sec:org1b0cbf0}

@@

\subsubsection{The effect of feedback}
\label{sec:org6038bfe}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgd2b8171}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orga7f95ff}

@@

\subsubsection{Study protocol}
\label{sec:org6624a8b}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:org1374951}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orga1e7779}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org93eda86}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgee42247}

@@

\subsection{Privacy}
\label{sec:org0ce2819}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org239dd78}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orga38c9fb}

@@

\section{Further Future Directions}
\label{sec:org2a60ef2}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org5aebc34}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org8f2e818}

@@

\subsection{Organic passwords}
\label{sec:orgd39940c}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org0118f24}

Where authenticity is nominally concerned with proving that you are who you say you are,
a less-frequently-asked question in the authentication literature is,
``are you really yourself?''
We all sometimes do or say regrettable things when we are feeling ``not quite ourselves,'' sometimes using devices or services with which we have authenticated ourself.
Can authentication ever verify not only your possession of your body, but of your ``right mind''?

A question raised earlier surrounds where passthoughts could still work if a person is drunk, having a migraine, or in distress (Section 3). 
Even if passthoughts fails when a user is in such an ``off-baseline'' state, 
passthoughts still may have utility (perhaps even \emph{added} utility) in certain authentication contexts.
For example, one may wish to allow themselves access to certain resources (e.g. bank accounts) when one's resting EEG state is not too much different from a pre-recorded baseline.

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org50e578e}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org9e676a3}

@@

\section{Conclusion}
\label{sec:org2301bf5}
